### 2017-06-01 R/TravelSurvey.Rmd
Update to Oregon Tutorial
and also TravelSurvey.Rmd

```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
foo <- unique(foo[,c("msa","msaname")])
msa <- foo$msa
msaname <- foo$msaname
names(msaname) <- msa

@@

```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
NHTS.df <- within(NHTS.df, {
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
    location <- msaname[location]
})
```

>>>>

```{r msa, eval = TRUE, fig.width = fig.width, echo = FALSE}
#foo <- read.csv(file ="./../data/cbsatocountycrosswalk.csv", colClasses = "character")
foo <- read.csv(file = "~/GHI/data/MSA.csv", colClasses = c("character","character"))
names(foo) <- c("msa","msaname")
#foo <- unique(foo[,c("msa","msaname")])
msa <- foo$msa
msaname <- foo$msaname
names(msaname) <- msa

@@

```{r TravelSurvey, eval = TRUE, fig.width = fig.width, echo = FALSE}
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
NHTS.df <- within(NHTS.df, {
    id <- paste0(houseID,"-",subjectID)
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
    location <- msaname[location]
})
```

### 2017-06-01 R/oregon.Rmd
Update to Oregon Tutorial
and also TravelSurvey.Rmd

# ITHIM Package Tutorial

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```
To begin using ITHIM one first must install it.  The package
_devtools_ makes this is easy.  Install _devtools_ using
`install.packages("devtools")` or some other way in _RStudio_.

@@

We use the function `install_github` to install the _master_ branch of
the _ITHIM_ package repository.

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("syounkin/ITHIM", ref="master")
library("ITHIM")
```

## Creating a Baseline ITHIM object

@@

For a description of the file format see the help page for
`createITHIM` using `help("createITHIM")`.

@@

* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)

```{r misc22, message = FALSE, eval = FALSE, warning = FALSE, error = TRUE, echo = FALSE}
purl("./tutorial.Rmd","./tutorialScript.R", documentation = 0)
```

#### Road Injuries File Format

@@

## Built-In Example

@@

### Session Info & Package Test
```{r sessionInfo, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("testthat")
test_package("ITHIM")
sessionInfo()
```

>>>>

# ITHIMr Demonstration for Portland, Oregon

```{r rm, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
rm(list=ls())
```

@@

The package _devtools_ makes installation of the _ITHIM_ package easy.
Install _devtools_ using `install.packages("devtools")` or some other
way in _RStudio_.  We use the function `install_github` from
_devtools_ to install the _devel_ branch of the _ITHIM_ package
repository.

```{r libs, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("devtools")
install_github("ITHIM/ITHIM", ref="devel")
```

```{r libs2, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("ITHIM")
library("tidyverse")
```

```{r figwidth, message = FALSE, eval = TRUE, warning = FALSE, error = TRUE, echo = FALSE}
fig.width <- 24 # inches
phi <- (1+sqrt(5))/2 # the golden ratio ;)
```

## Create a Baseline ITHIM object

@@

The file_activeTransprtTime.csv_ contains weekly mean travel times in
minutes, stratified by age, sex and mode (walk,cycle). Currently eight
age classes are used for the _ITHIM_ model.  For a more detailed
description of the file format see the help page for `createITHIM`
using `help("createITHIM")`.

@@

Sample Files
* [activeTransportTime.csv](https://github.com/syounkin/ITHIM/blob/master/inst/activeTransportTime.csv)
* [gbd.csv](https://github.com/syounkin/ITHIM/blob/master/inst/gbd.csv)

@@

## Active Travel Estimates for Portland (MSA 6442)

We use the NAtional Household Transportation Survey data stored as a _TravelSurvey_ R object to estimate mean walk and cycle times.

```{r oregon23, eval = TRUE, echo = FALSE, message = FALSE, warnings = FALSE, results = "hide", fig.width = fig.width, fig.asp = 1/phi}
install_github("syounkin/TravelSurvey", ref="master")
library("TravelSurvey")
NHTS.ts <- readRDS(file = "~/GHI/R/data/NHTS.TravelSurvey.rds")
NHTS.df <- dplyr::left_join(NHTS.ts@person, NHTS.ts@trip, by = c("houseID","subjectID"))
```

```{r oregon333, eval = TRUE, echo = FALSE, fig.width = fig.width, fig.asp = 1/phi}
NHTS.df <- within(NHTS.df, {
    id <- paste0(houseID,"-",subjectID)
    duration <- ifelse(is.na(duration), 0, duration)
    mode <- factor(mode, levels = c("walk","cycle","other"))
    age <- factor(age, levels = c("00-04","05-14","15-29","30-44","45-59","60-69","70-79","80+"))
    sex <- factor(sex, levels = c("M","F"))
})
```

```{r oregon2, eval = TRUE, echo = TRUE, fig.width = fig.width, fig.asp = 1/phi}
sampleSize <- NHTS.df %>% filter(location == "6442") %>% group_by(sex,age) %>% summarise(n=n_distinct(id))

tripDuration <- NHTS.df %>% filter(location == "6442") %>% group_by(mode,sex,age) %>% summarise(total=sum(duration))

activeTravelOregon <- right_join(sampleSize,tripDuration, by = c("sex","age")) %>% ungroup() %>% mutate(mean = total/n*7) %>% select(mode,sex,age,mean) %>% complete(mode,sex,age) %>% filter(mode %in% c("walk","cycle")) %>% as.data.frame()

activeTravelOregon
```

### Built-In Example

@@

### Road Injuries
#### Road Injuries File Format

### Session Info & Package Test
```{r sessionInfo, eval = FALSE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
library("testthat")
test_package("ITHIM")
sessionInfo()
```
